---
title: "Cross validation using environmental blocking"
subtitle: "Based on annual fire data"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
params:
  test_run: FALSE
  s: '_annf3_A-P_entire'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

The goal here is to split the data into multiple training/testing
blocks based on clustering grid-cells by climate.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE,
                      message = FALSE)
```


# Dependencies


```{r, message = FALSE, warning = FALSE}
library(blockCV) # for creating environmental blocks
library(sf)
library(knitr)
library(tidyverse)
library(stars)
library(terra)
source("src/modeling_functions.R")
source("src/general_functions.R")
source("src/basemaps.R")
source("src/fig_params.R")
theme_set(theme_classic())
```

params

```{r}
print(params)
test_run <- params$test_run # runs code on data subset, for testing
s <- params$s

n_pdp <- 1e5
```


## Read in data

Read in model object (for the formula)

```{r}
path_mod <- paste0("models/glm_binomial_models_v3", s, ".RDS")
mod1 <- butcher::butcher(readRDS(path_mod)) # to save memory
```

Reading in predictor vars

```{r}
df_ann1 <- read_main_csv() %>% 
  # removing excess rows for memory saving
  select(-matches('burn_frac'), -matches('nfire_cwf_centroid'),
         -matches('weight'))

df_ann2 <- df_ann1 %>% 
  select(cell_num, pfgAGB, afgAGB, nfire_cwf, cwf_prop, MAP, MAT, prcpPropSum)
```

Template (cell number) raster

```{r}
template <- rast("data_processed/data_publication/cell_number_ids.tif")
```



# functions

```{r}
# summarizing across columns of interest
summarize_pred_response <- function(df) {
  out <- df %>% 
    mutate(MAT = MAT - 273.15,
           # convert from probability to percent
         across(matches("cwf_prop"),
                        .fns = function(x) x*100)) %>% 
    summarise(across(c(MAT, MAP, prcpPropSum, afgAGB, pfgAGB, matches("cwf_prop")),
                  .fns = mean))
  out
}
```


# Determine Blocks


Combine climate data into a single raster for use in blockCV::envBlock(),
and create an sf object of fake datapoints so that function will work. 

```{r blocks}
# average climate conditions across all the years 
df_clim_avg <- df_ann1 %>% 
  summarize_yearly(weighted_vars = c("MAT", "MAP", "prcpPropSum"),
                   sum_vars = NULL)

r_clim2 <- fill_raster(df_clim_avg, template)
template <- mask(template, r_clim2[[1]]) # in case template has additional cells in it
n <- ncell(r_clim2)
cell_coords <- xyFromCell(template, 1:n) %>% 
  as.data.frame() %>% 
  mutate(
    cell_num = as.numeric(values(template))) %>% 
  filter(!is.na(cell_num)) %>% 
  # sf object of the center of sagebrush biome grid points
  st_as_sf(coords = c('x', 'y'), crs = crs(template))


k <- 5
fold_name <- paste0('fold', 1:k)

set.seed(124) # envBlock uses kmeans which is affected by the seed

# environmental blocks
eb1 <- blockCV::cv_cluster(x = cell_coords, # coords of center of eath grid cell
                           r = r_clim2,
                           k = k,
                           scale = TRUE,
                           raster_cluster = FALSE)


# checking that size of all test datasets sums to the 
# number of grid cells in the original dataset
cell_nums <- unique(df_clim_avg$cell_num)
stopifnot(length(unique(cell_nums)) 
          == sum(eb1$records$train[1] + eb1$records$test[1]))
```

# Prepare Data


```{r}
pred_vars <- c("afgAGB", "pfgAGB", "MAT", "MAP", "prcpPropSum")
clim_vars <- c("MAT", "MAP", "prcpPropSum")
names(pred_vars) <- pred_vars
response_vars <- c("cwf_prop", "cwf_prop_pred")

```


## Split into train/test

```{r}
fold_endices <- cell_coords
fold_endices$geometry <- NULL
fold_endices$fold <- eb1$folds_ids
fold_endices <- as_tibble(fold_endices) %>% 
  # renumbering so number (and colors) will
  # (more or less) match with the first submission
  # to fire ecology (for easier interpretation)
  mutate(fold = c(1, 2, 4, 3, 5)[fold]) %>% 
  arrange(fold, cell_num)
test_cell_nums <- fold_endices %>% 
  mutate(fold = paste0('fold', fold)) %>% 
  split(., .$fold)

# data for a given fold
dfs_test1 <- map(test_cell_nums, function(x) {
  df_ann2[df_ann2$cell_num %in% x$cell_num, ] 
})

# data for the other folds
dfs_train1 <- map(test_cell_nums, function(x) {
  df_ann2[!df_ann2$cell_num %in% x$cell_num, ] 
})

```

```{r}
set.seed(111)
if (test_run) {
  n <- 1000
  dfs_test1 <- map(dfs_test1, ungroup) %>% 
    map(slice_sample, n = n)
  dfs_train1 <- map(dfs_train1, ungroup) %>% 
    map(slice_sample, n = n)
  df_ann3 <- slice_sample(ungroup(df_ann2), n = n*5)
} else {
  df_ann3 <- df_ann2
}

# small dataset used for creating pdp plots (so code isn't too slow)
train_pdp <- df_ann3 %>% 
  slice_sample(n = n_pdp) %>% 
  select(all_of(pred_vars))

dfs_test_pdp1 <- map(dfs_test1, \(x) slice_sample(x[pred_vars], n = n_pdp)) # samples of the test datasets (for making pdp plots)
dfs_train_pdp1 <- map(dfs_train1, \(x) slice_sample(x[pred_vars], n = n_pdp)) # samples of the training datasets (for making pdp plots)
```


# Describe folds



```{r}
cols_fold <- RColorBrewer::brewer.pal(k, 'RdYlBu')
r_folds1 <- fill_raster(fold_endices, template)

writeRaster(r_folds1, 'data_processed/cv_folds_v2.tif', overwrite = TRUE)

# the base and bbox objects come from src/basemaps.R
g <- ggplot() +
  geom_stars(data = st_as_stars(r_folds1), aes(fill = factor(fold, levels = 1:k))) +
  basemap_g() +
  scale_fill_manual(values = cols_fold,
                     na.value ='white',
                    name = 'Fold',
                    na.translate = FALSE) 
g

if(!test_run){
png("figures/cv/folds_map_v1.png",width = 5, height = 4, units = 'in',
       res = 600)
g
dev.off()
}
```


# Fit models

All models have the same formula. But are fit to data where one
block of data is held out. (i.e. each model fit to ~4/5 of the data)

```{r fit_models}
form1 <- as.formula(mod1$formula)

# mods fit to 4/5 of the data
mods_train1 <- map(dfs_train1, function(df) {
  butcher::butcher(glm(form1, family = 'binomial', data = df))
})


# predictions on the test data set (1/5 of the data)
dfs_test2 <- map2(mods_train1, dfs_test1, function(mod, df) {
  df$cwf_prop_pred <- predict(mod, newdata = df, type = 'response')
  df
})

```


## PDP

These are true PDP plots--an earlier version of this code created
'pseudo' pdp plots where I was predicting across the range of a given predictor
variable, with other variables held at their mean. 


```{r }
lookup_var <- var2lab(x = NULL, units_md = TRUE)

mods_train2 <- c(mods_train1, list('biome-wide' = mod1))
dfs_train_pdp2 <- c(dfs_train_pdp1, 
                    list('biome-wide' = slice_sample(df_ann3[pred_vars],
                                                     n = n_pdp)))

letter_df <- tibble(
  letter = fig_letters[1:length(lookup_var)],
  pred_var_name = factor(lookup_var),
  value = -Inf,
  yhat = Inf
)

```

```{r create_pdp_df, cache = FALSE, eval= !test_run}
df_pdp_by_fold1 <- map2_dfr(mods_train2, dfs_train_pdp2, function(mod, df) {
  create_pdp_df(mod = mod, mod_vars = names(lookup_var), data = df)
},
.id = 'fold')

```


```{r fig.height = 3.5, fig.width = 6.8, eval = !test_run}
cols_fold2 <- cols_fold
names(cols_fold2) <- fold_name
df_pdp_by_fold2 <- df_pdp_by_fold1 %>% 
  mutate(pred_var_name = var2lab(pred_var, units_md = TRUE))

g2 <- ggplot(df_pdp_by_fold2, aes(value, yhat*100)) +
  geom_line(aes(color = fold)) +
  geom_text(data = letter_df, aes(label = letter),
            hjust = -0.8,
            vjust = 1) +
  facet_wrap(~pred_var_name, scales = 'free', strip.position = "bottom") +
  theme(strip.text = ggtext::element_markdown(),
      strip.placement = "outside",
      axis.title.x = element_blank(),
      strip.background = element_blank()) +
    expand_limits(y = 3) +
  labs(y = lab_fireProbPerc)+
  scale_color_manual(values = c(cols_fold2, 'biome-wide' = 'black'), 
                     name = "Fold left out")
g2
if(!test_run){
jpeg(paste0("figures/pdp/cv_pdp_v1", s, ".jpeg"), units = "in", res = 600,
     width = 6.8, height = 4)
  g2
dev.off()
}
```

## Summary stats

Summarizing fire occurrence by fold. And predictions are made by the main model. 

```{r}
df_ann3_fold <- df_ann3 %>% 
  left_join(bind_rows(test_cell_nums), by = 'cell_num')

df_ann3_fold <- df_ann3_fold %>% 
  split(., .$fold) %>%
  # cross validation predictions
  # for each fold creat prediction from model not fit to data from that fold
  map2(mods_train1, function(df, mod) {
    df$cwf_prop_pred_cv <- predict(mod, newdata = df,
                                   type = 'response')
    df
  }) %>% 
  bind_rows(.id = "fold")

# prediction from the main model
df_ann3_fold$cwf_prop_pred_main <- predict(mod1, newdata = df_ann3_fold,
                                 type = 'response')

digits <- c(NA, 1, 0, 3, 1, 1, 2, 2, 2)

df_ann3_fold %>% 
  group_by(fold) %>% 
  summarize_pred_response() %>% 
  kable(caption = "Mean pred vars and observed/predicted fire probability by fold",
        digits = digits) 

df_ann3_fold %>% 
  ungroup() %>% 
  summarize_pred_response() %>% 
  kable(caption = "Mean pred vars and observed/predicted fire probability across biome",
        digits = digits[-1]) 
```



## size

Note here the number of observations per fold means the number of rows
of data (not # of pixels). Some pixels have multiple observations (i.e.,
when a fire occurred)

```{r}
# number of grid cells in the fold ( raw data)
df_ann3_fold %>% 
  group_by(fold) %>% 
  summarize(fold_size = n()) %>% 
  kable(caption = "Number of observations per fold")

```


## Observed vs predicted

### Deciles

```{r quantile_plots, fig.width = 5, fig.height = 3.5 }

pred_test_deciles1 <- map(dfs_test2, predvars2deciles, pred_vars = pred_vars,
                 response_vars = response_vars)

map2(pred_test_deciles1, names(pred_test_deciles1), function(df, fold) {
  decile_dotplot_pq(df = df) +
    labs(subtitle = paste0("Predictions on ", fold, 
                           " (model fit to other folds)")) +
    coord_cartesian(ylim = c(0, 3.5))
})
```




### deciles filtered

Predictions are made by models fit to the other folds

```{r filtered_quantile}
pred_test_deciles_filt <- map(dfs_test2, predvars2deciles, pred_vars = pred_vars,
                 response_vars = response_vars,
                 filter_var = TRUE)
```

```{r fig.width = 7, fig.height = 6 }

g_filt <- map2(pred_test_deciles_filt, names(pred_test_deciles_filt), function(df, fold) {
  g <- decile_dotplot_filtered_pq2(df = df) +
    patchwork::plot_annotation(subtitle = paste0("Predictions on ", fold, 
                           " (model fit to other folds)"))
  
  if(!test_run){
  jpeg(paste0("figures/quantile_plots/cv_quantile_plots_filt_", fold, s, 
              ".jpeg"), units = "in", res = 600,width = 7, height = 6 )
    print(g) 
  dev.off()
  }
  g
})

g_filt
```

### deciles filtered (veg)

Predictions for each fold made by the main model (i.e. this part isn't actually
cross validation)

```{r fig.height = 10, fig.width = 5}

pred_test_deciles_filt_main <- df_ann3_fold %>%
  rename(cwf_prop_pred = cwf_prop_pred_main) %>% 
  split(., .$fold) %>% 
  map(predvars2deciles, pred_vars = pred_vars,
                 response_vars = response_vars,
                 filter_var = TRUE,
                 filter_vars = pred_vars)

figs <- map(pred_test_deciles_filt_main, decile_dotplot_filtered_pq)

map2(figs, names(figs), function(g, fold) {
  g +
    labs(subtitle = paste('Balanced data from', fold, 
                          '(predictions by main model)'))
})
```

### deciles filtered (clim)

```{r fig.height = 10, fig.width = 7}
figs2 <- map(pred_test_deciles_filt_main, decile_dotplot_filtered_pq,
    xvars = clim_vars)

map2(figs2, names(figs2), function(g, fold) {
  g +
    labs(subtitle = paste('Balanced data from', fold, 
                          '(predictions by main model)'),
         caption = 'y limits constrained') +
    coord_cartesian(ylim = c(0, 4))
})
```



# Fitting model to each fold
Here fitting data 1/5 of the data instead of 4/5
Fitting separate models to each fold, for computational reasons not 
going through model selection. See earlier versions of this script
to get code on model selection at each fold

# model fitting

```{r one_mod_per_fold}
# mods fit to 4/5 of the data
mods_test1 <- map(dfs_test1, function(df) {
  butcher::butcher(glm(form1, family = 'binomial', data = df))
})

```


## PDP

Creating partial dependence plots of the model fit to each fold

```{r pdp_best, eval = FALSE}

df_pdp_mod_fold1 <- map2_dfr(mods_test1, dfs_test_pdp1, function(mod, df) {
  create_pdp_df(mod = mod, mod_vars = names(lookup_var), data = df)
},
.id = 'fold')

```

```{r fig.height = 3.5, fig.width = 6.8, eval = FALSE}
df_pdp_mod_fold1 <- df_pdp_mod_fold1 %>% 
  mutate(pred_var_name = var2lab(pred_var, units_md = TRUE))

g2 <- ggplot(df_pdp_mod_fold1, aes(value, yhat*100)) +
  geom_line(aes(color = fold)) +
  facet_wrap(~pred_var_name, scales = 'free', strip.position = "bottom") +
  theme(strip.text = ggtext::element_markdown(),
      strip.placement = "outside",
      axis.title.x = element_blank(),
      strip.background = element_blank()) +
    expand_limits(y = 3) +
  labs(y = lab_fireProbPerc)+
  scale_color_manual(values = cols_fold, 
                     name = "Fold model fit to")

g2

g2 +
  coord_cartesian(ylim = c(0, 3))
```

# session info

Hash of current commit (i.e. to ID the version of the code used)

```{r}
system("git rev-parse HEAD", intern=TRUE)
```

Packages etc.

```{r}
sessionInfo()

```


