---
title: "Models to convert RAP shrub cover to biomass"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

# Overview

The rangeland analysis platform provides shrub cover but not shrub biomass.
Therefore the models we developed to predict fire probability use shrub cover. 
However, stepwat2 output is shrub biomass, therefore the goal here is to find
a relationship between remotely sensed shrub cover and stepwat2 simulated biomass.

Additionally here I will look at the relationship between RAP annual and perennial
biomass and the respective biomass simulated by stepwat2--as an exercise
in assessing the agreement between these very much independent data sources.

I will be using RAP data that 1) falls within the biome mask, and 2) falls
within the mask of the stepwat2 upscaling. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r dependencies, message = FALSE, warning=FALSE}
library(tidyverse)
library(terra)
theme_set(theme_classic())
```

# Load data

RAP data

```{r}
r_rap1 <- rast("data_processed/RAP/RAP_afgAGB-pfgAGB-shrCover_1984-2019_med-max_1000m_sw2sim-extent_v1.tif")
names(r_rap1)


# grabbing this to use as a mask for ths sw2sim-extent data
r_rap_biome1 <- rast("data_processed/RAP/RAP_afgAGB-pfgAGB-shrCover_1984-2019_median_1000m_sagebrush-biome-mask_v1.tif")
```

STEPWAT2 data

```{r}
r_bio1 <- rast("../grazing_effects/data_processed/interpolated_rasters/bio_future_median_across_GCMs.tif")

# layers of interest:
lyrs_perennial <- paste0("c4on_", 
                         c("C3Pgrass", "C4Pgrass", "Pforb"),
                         "_biomass_Current_Current_Light")

lyrs_annual <- paste0("c4on_", 
                      c("Cheatgrass", "Aforb"),
                      "_biomass_Current_Current_Light")

lyrs_shrub <- "c4on_Shrub_biomass_Current_Current_Light"

r_bio1 <- r_bio1[[c(lyrs_perennial, lyrs_annual, lyrs_shrub)]]
```

SEI data

Using this data to create a mask of only core areas

```{r}
r_c9 <- rast("../SEI/data_processed/transitions/SEIv11_9ClassTransition_2017_2020_90_ClimateOnly_RCP85_2030-2060_median.tif")
r_c9 <- r_c9[[1]]
```

# Data Prep

Extract core areas. In this rasters 1 is stable core, 2, is core becomes grow,
and 3 is core becomes other. so 1, 2, and 3, together make up 'current core'

```{r}
r_core1 <- r_c9
# values from 1 to 3 will become 1, others become NA
r_core2 <- classify(r_core1, 
                    rcl =matrix(c(0.5, 3.5, 1), byrow = TRUE, nrow = 1),
                    othersNA = TRUE)

# core areas are 1, growth areas are 2
r_core_grow1 <- classify(r_core1, 
                    rcl =matrix(c(0.5, 3.5, 1, 3.5, 6.5, 2, 6.5, 9.5, 3), byrow = TRUE, nrow = 3),
                    othersNA = TRUE)
plot(r_core2)
```


Makes extents equal, and mask out cells that
are not part of the sagebrush-biome mask.


```{r fix_extent}
mask_biome1 <- r_rap_biome1[[1]]

# r_rap_biome1 has slightly different resolution than the other 
# rasters, so resampling to the same resolution
# Note--this resampling does change which locations are masked out--slightly
mask_biome2 <- resample(mask_biome1, r_bio1)
r_core3 <- resample(r_core2, r_bio1)
r_bio1 <- crop(r_bio1, mask_biome2)

r_rap2 <- crop(r_rap1, mask_biome2)

r_rap2 <- mask(r_rap2, mask = mask_biome2)
r_bio2 <- mask(r_bio1, mask_biome2)
```


Calculate afgAGB and pfgAGB from stepwat output

```{r stepwat_sim_df}

# simulated biomass
df1 <- tibble(
    # annuals
  afgAGBsw2 = sum(r_bio2[[lyrs_annual]]) %>% 
    values() %>% 
    as.vector(),
  # perennials
  pfgAGBsw2 = pfgAGB <- sum(r_bio2[[lyrs_perennial]]) %>% 
    values() %>% 
    as.vector(),
  # shrubs
  shrAGBsw2 = r_bio2[[lyrs_shrub]] %>% 
    values() %>% 
    as.vector(),
  core = r_core3[[1]] %>% 
    values() %>% 
    as.vector() 
  )
df1$core <- ifelse(is.na(df1$core), 'Note core', 'Core')
```

Add RAP data to dataframe

```{r}
rap_lyrs1 <- names(r_rap1) 
rap_lyrs2 <- paste0(rap_lyrs1, "_rap")

rap_lyrs1

for(i in seq_along(rap_lyrs1)) {
  df1[[rap_lyrs2[i]]] <-  r_rap2[[rap_lyrs1[[i]]]] %>% 
    values() %>% 
    as.vector()
}

df2 <- drop_na(df1)
remove('df1')

```
## biome mask--for AGB by SEI

Compiling RAP data, with SEI categorizating for the given pixel. 
Note this is doen at 1km resolution, so some inaccuracies will exist. 

```{r}

# r_rap_biome1 has slightly different resolution than the other 
# rasters, so resampling to the same resolution
# Note--this resampling does change which locations are masked out--slightly
r_core_grow1 <- resample(r_core_grow1, r_rap_biome1)

r_core_grow2 <- crop(r_core_grow1, r_rap_biome1)

df_agb_sei1 <- tibble(
  SEI = as.vector(values(r_core_grow2)),
  afgAGB = as.vector(values(r_rap_biome1[['afgAGB']])),
  pfgAGB = as.vector(values(r_rap_biome1[['pfgAGB']]))
)

df_agb_sei2 <- df_agb_sei1 %>% 
  mutate(SEI = case_when(SEI == 1 ~ "Core",
                         SEI == 2 ~ "Growth",
                         SEI == 3 ~ "Other")) %>% 
  filter(!is.na(SEI))
```

# Histograms (RAP by SEI)

```{r}
g <- ggplot(df_agb_sei2) +
  facet_wrap(~SEI, nrow = 2, scales = 'free_y') +
  labs(caption = "RAP data across the Doherty et al. sagebrush biome mask,
       grouped by SEI classification")

g + 
  geom_histogram(aes(x = afgAGB), bins = 200) + 
  coord_cartesian(xlim = c(0, 50))
g + 
  geom_histogram(aes(x = pfgAGB), bins = 200) + 
  coord_cartesian(xlim = c(0, 150))

```



# Exploratory Analysis--RAP vs SW2

```{r message = FALSE}
df3 <- df2 %>% 
  filter(shrAGBsw2 < 740)
names(df2)
# variables to compare (vector not used in the code at the moment)
comparisons <- c(
  "afgAGBMax_rap" = "afgAGBsw2",
  "afgAGBMed_rap" = "afgAGBsw2",
  "pfgAGBMax_rap" = "pfgAGBsw2",
  "pfgAGBMed_rap" = "pfgAGBsw2",
  "shrCoverMax_rap" = "shrAGBsw2",
  "shrCoverMed_rap" = "shrAGBsw2"
)

df3 %>% 
  ggplot( aes(shrAGBsw2, shrCoverMax_rap)) +
    geom_point(size = 0.2, alpha = 0.1) +
    geom_smooth(se= FALSE) +
    geom_smooth(method = 'lm', se = FALSE) +
    labs(x = "Shrub biomass (STEPWAT2)",
         y = "Max shrub cover (%; RAP)",
         subtitle = 'Max shrub cover') +
    facet_wrap(~core)

df3 %>% 
  ggplot( aes(shrAGBsw2, shrCoverMed_rap)) +
    geom_point(size = 0.2, alpha = 0.1) +
    geom_smooth(se= FALSE) +
    geom_smooth(method = 'lm', se = FALSE) +
    labs(x = "Shrub biomass (STEPWAT2)",
         y = "Median shrub cover (%; RAP)",
         subtitle = 'Median shrub cover') +
    facet_wrap(~core)

```

## percentiles

Grouping stepwat2 shrub biomass by unique value (i.e. by plot), and
then taking the 95th percentile of 
```{r warning = FALSE, message = FALSE}
df3_pcent <- df3 %>% 
  group_by(shrAGBsw2) %>% 
  summarise(across(c("shrCoverMax_rap", "shrCoverMed_rap"), 
                   .fns = quantile, probs = 0.9),
            .groups = 'drop')

ggplot(df3_pcent, aes(shrAGBsw2, shrCoverMed_rap)) +
  geom_point() +
  geom_smooth(method = 'glm') +
#  facet_wrap(~core) +
  labs(subtitle = '90th pcentile of shrub cover (RAP) across pixels, for each level of shrub biomass (SW2)')
  
  
```

# Models

