---
title: "Models to convert RAP shrub cover to biomass"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

# Overview

The rangeland analysis platform provides shrub cover but not shrub biomass.
Therefore the models we developed to predict fire probability use shrub cover. 
However, stepwat2 output is shrub biomass, therefore the goal here is to find
a relationship between remotely sensed shrub cover and stepwat2 simulated biomass.

Additionally here I will look at the relationship between RAP annual and perennial
biomass and the respective biomass simulated by stepwat2--as an exercise
in assessing the agreement between these very much independent data sources.

I will be using RAP data that 1) falls within the biome mask, and 2) falls
within the mask of the stepwat2 upscaling. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r dependencies, message = FALSE, warning=FALSE}
library(tidyverse)
library(terra)
theme_set(theme_classic())
```

# Load data

RAP data

```{r}
r_rap1 <- rast("data_processed/RAP/RAP_afgAGB-pfgAGB-shrCover_1984-2019_med-max_1000m_sw2sim-extent_v1.tif")
names(r_rap1)

# grabbing this to use as a mask for ths sw2sim-extent data
r_rap_biome1 <- rast("data_processed/RAP/RAP_afgAGB-pfgAGB-shrCover_1984-2019_median_1000m_sagebrush-biome-mask_v1.tif")
```

STEPWAT2 data

```{r}
r_bio1 <- rast("../grazing_effects/data_processed/interpolated_rasters/bio_future_median_across_GCMs.tif")

# layers of interest:
lyrs_perennial <- paste0("c4on_", 
                         c("C3Pgrass", "C4Pgrass", "Pforb"),
                         "_biomass_Current_Current_Light")

lyrs_annual <- paste0("c4on_", 
                      c("Cheatgrass", "Aforb"),
                      "_biomass_Current_Current_Light")

lyrs_shrub <- "c4on_Shrub_biomass_Current_Current_Light"

r_bio1 <- r_bio1[[c(lyrs_perennial, lyrs_annual, lyrs_shrub)]]
```

# Data Prep

Makes extents equal, and mask out cells that
are not part of the sagebrush-biome mask.



```{r fix_extent}
mask_biome1 <- r_rap_biome1[[1]]

# r_rap_biome1 has slightly different resolution than the other 
# rasters, so resampling to the same resolution
# Note--this resampling does change which locations are masked out--slightly
mask_biome2 <- resample(mask_biome1, r_bio1)

r_bio1 <- crop(r_bio1, mask_biome2)

r_rap2 <- crop(r_rap1, mask_biome2)

r_rap2 <- mask(r_rap2, mask = mask_biome2)
r_bio2 <- mask(r_bio1, mask_biome2)

```


Calculate afgAGB and pfgAGB from stepwat output

```{r stepwat_sim_df}

# simulated biomass
df1 <- tibble(
    # annuals
  afgAGBsw2 = sum(r_bio2[[lyrs_annual]]) %>% 
    values() %>% 
    as.vector(),
  # perennials
  pfgAGBsw2 = pfgAGB <- sum(r_bio2[[lyrs_perennial]]) %>% 
    values() %>% 
    as.vector(),
  # shrubs
  shrAGBsw2 = r_bio2[[lyrs_shrub]] %>% 
    values() %>% 
    as.vector()
)

```

Add RAP data to dataframe

```{r}
rap_lyrs1 <- names(r_rap1) 
rap_lyrs2 <- paste0(rap_lyrs1, "_rap")

rap_lyrs1

for(i in seq_along(rap_lyrs1)) {
  df1[[rap_lyrs2[i]]] <-  r_rap2[[rap_lyrs1[[i]]]] %>% 
    values() %>% 
    as.vector()
}

df2 <- drop_na(df1)
remove('df1')

```

# Exploratory Analysis

```{r}

names(df2)
# variables to compare (vector not used in the code at the moment)
comparisons <- c(
  "afgAGBMax_rap" = "afgAGBsw2",
  "afgAGBMed_rap" = "afgAGBsw2",
  "pfgAGBMax_rap" = "pfgAGBsw2",
  "pfgAGBMed_rap" = "pfgAGBsw2",
  "shrCoverMax_rap" = "shrAGBsw2",
  "shrCoverMed_rap" = "shrAGBsw2"
)

ggplot(df2, aes(shrAGBsw2, shrCoverMax_rap)) +
  geom_point(size = 0.2, alpha = 0.1) +
  geom_smooth(se= FALSE) +
  geom_smooth(method = 'lm', se = FALSE) +
  labs(x = "Shrub biomass (STEPWAT2)",
       y = "Max shrub cover (%; RAP)")


```


# Models

