---
title: "Models of fire probability for the sagebrush biome"
subtitle: "Resampled/stratified data"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

Here the training data is resampled to be more balanced (i.e., upsampling
bins of rare data, downsampling bins of common data). 

Note a concern with the up-sampling is that many bins start with only one
datapoint in them, which is then repeated many times in the resampled datset.
Apparently, having duplicated points like this can increase the risk of 
over-fitting. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, cache.lazy = FALSE)
```

# Dependencies 

Sourcing the script that creates a dataframe of historical
fore occurrence, as well as biotic and climate
predictors for those same grid-cells. 

```{r source_data, warning=FALSE, message=FALSE, cache = TRUE}
source("scripts/04_create_biome-mask_dataframe.R")
```

```{r warning=FALSE, message=FALSE}
source("src/general_functions.R")
source("src/fig_params.R")
source("src/modeling_functions.R")
source("src/resample_functions.R")
library(randomForest)
library(tidyverse)
library(margins) # for cplot()
library(GGally) # for ggpairs()
library(gnm) # generalized non-linear models
library(pdp) # for partial dependence plots
library(splines) # for ns() (splines) inside formulas of glm()
library(gridExtra)
library(knitr)
theme_set(theme_classic())
```


# Prep data

```{r prep_data, cache=TRUE}

# converting a fire count into a factor (true if at least 1 fire occured)
occur_factor <- function(nfire){
  factor(nfire > 0, levels = c("TRUE", "FALSE"))
}
dfs_biome3 <- map(dfs_biome2, function(df) {
  out <- df %>% 
  # probability of fire in a given year
    mutate(# number of years the mtbs count data corresponds to (for binomial glm)
           # all count datasets (mtbs and ifph) are 35 years (change if updated)
           mtbs_n = 36, 
           # proportion of years with fires
           mtbs_prop = nfire_mtbs/mtbs_n,
           ifph_prop = nfire_ifph/mtbs_n,
           comb_prop = nfire_comb/mtbs_n,
           lba_prop = nfire_lba/mtbs_n
           ) %>% 
    # creating true/false occurrence cols for each datasets
    mutate(across(matches("^nfire_"), .fns = list(occur = occur_factor),
                  .names = "{.fn}_{.col}")) %>% 
    rename_with(.fn = str_replace, .cols = matches("^occur_"),
                pattern = "_nfire", replacement = "")
})

method <- names(dfs_biome3)
names(method) <- method

```

```{r}
set.seed(1234)
pred_vars <- c("afgAGB", "pfgAGB", "MAT", "MAP", "prcpPropSum")
names(pred_vars) <- pred_vars
# so that herbAGB is in the final dataframes
pred_vars2 <- unique(c(pred_vars, 'herbAGB'))

# predictor vars are the same in both dfs
df_pred <- dfs_biome3[[1]][, pred_vars2]
```


## Bin data

```{r}
df_pred_bin1 <- bin_df(df_pred, cols = pred_vars)
head(df_pred_bin1)
```

## Explore bins

Number of unique bins

```{r}
n_bins <- df_pred_bin1$bin_all %>% 
  unique %>% 
  length()
n_bins
```

The 10 bins with the most 

```{r}
n_per_bin_table <- table(df_pred_bin1$bin_all)

n_per_bin <- as.vector(n_per_bin_table)

summary(n_per_bin)
names(n_per_bin) <- dimnames(n_per_bin_table)[[1]]


sort(n_per_bin, decreasing = TRUE)[1:10]
```

Histograms of the number of observations per bin

```{r }
  hist(n_per_bin, breaks = 200,
       main = 'number of observations in each multi-dimensional bin',
       xlab = "Number of observations in bin")

  hist(n_per_bin[n_per_bin < 100], breaks = 100,
       main = 'number of observations in each multi-dimensional bin \n(xlim restricted)',
       xlab = "Number of observations in bin")
```

## Resample data

```{r resample, message = FALSE}
# note this chunk of code can't be safely re-run without
# rerunning code above that creates dfs_biome3

df_pred_resample1 <-  df_pred_bin1 %>% 
  mutate(row_num = 1:nrow(.)) %>% 
  group_by(bin_all) %>% 
  slice_sample(
  # making each bin the same size so end up with
  # ~ as many rows as the original dataset
  n = round(nrow(df_pred_bin1)/n_bins),
  replace = TRUE)

# repeating the same rows in both dataframes in dfs_biome3
dfs_biome3 <- map(dfs_biome3, function(df) {
  df$bin_all <- df_pred_bin1$bin_all
  
  out <- df[df_pred_resample1$row_num, ]
  out
})

# checking that all bins have the same number of observations
dfs_biome3$paint$bin_all %>% 
  table() %>% 
  as.vector() %>% 
  unique()
```


## training data

Note--previously a small sample for now to make model fitting quicker.
Now the training dataset is all the data.

```{r train, dependson='prep_data', cache = TRUE}


# n = 10000 # for testing
n = nrow(dfs_biome3[[1]]) # fit models on all the data
         
row_nums <- 1:nrow(dfs_biome3[[1]])
rows_train <- sample(row_nums, size = n, replace = FALSE)

# rows for test dataset
# note set up this way, in case the training
# dataset is all the data
rows2sample <- if(n == length(rows_train)) {
  row_nums
} else {
  row_nums[!row_nums %in% rows_train]
}

# rows used in testing dataset
rows_test <- sample(rows2sample, size = 40000, replace = FALSE)

# selecting rows this was so both dataframes in list have same rows
# and differences in models will only be b/ of differences in response

dfs_train <- map(dfs_biome3, function(df) as.data.frame(df[rows_train, ]))
dfs_test <- map(dfs_biome3, function(df) as.data.frame(df[rows_test, ]))

# all variables (transformed or not, that could be predictor vars)
all_pred_vars <- names(dfs_biome3[[1]]) %>% 
  str_subset(pattern = paste0("(", 
                              paste0(pred_vars2, collapse = ")|("), 
                              ")")
             )
train_pred <- dfs_train[[1]][, all_pred_vars]
test_pred <- dfs_test[[1]][, all_pred_vars]
```

# Exploratory figs & summary values


```{r summary_table}

create_summary <- function(df) {
  df %>% 
    pivot_longer(cols = everything(),
                 names_to = 'variable') %>% 
    group_by(variable) %>% 
    summarise(across(value, .fns = list(mean = mean, min = min, 
                                        median = median, max = max))) %>% 
    mutate(across(where(is.numeric), round, 4))
}

dfs_biome3[[1]][pred_vars2] %>% 
  create_summary() %>% 
  knitr::kable(caption = 'summaries of predictor variables')


response_summary <- map(dfs_biome3, function(df) {
  df %>% 
    dplyr::select(where(is.numeric), -all_of(pred_vars2), -matches("^occur_")) %>% 
    create_summary()
})

kable(response_summary$paint, 
      caption = 'summaries of response variables, calculated using paint')

kable(response_summary$reduceToImage, 
      caption = 'summaries of response variables, calculated using reduceToImage method')

```

## Plot predictor vars against each other

here using pred dataframe, because smaller and this code is slow. 

```{r pred_v_pred, cache=TRUE, dependson='train'}
ggpairs(test_pred %>% select(-matches("_")),
        lower = list(continuous = GGally::wrap("points", alpha = 0.1,    size=0.2)))
```

## boxplots-- # of fires vs predictor variables

```{r n_fires_boxplots, fig.height=9, fig.width=8, cache = TRUE, dependson='prep_data'}

# vectors of names of response variables
vars_nfire <- names(dfs_biome3[[1]]) %>% 
  str_subset("^nfire_")

vars_occur <- names(dfs_biome3[[1]]) %>% 
  str_subset("^occur_")

names(vars_occur) <- str_replace(vars_occur, "occur_", "")

vars_prop <- str_replace(vars_occur, "occur_", "") %>% 
  paste0(., "_prop")

names(vars_prop) <- names(vars_occur)

# longformat dataframes for making boxplots
dfs_biome_long <- map(dfs_biome3, function(df) {
  # for some reason select() was giving me problems
  predvars2long(df, response_vars = c(vars_nfire, vars_prop), 
                pred_vars = pred_vars) %>% 
    mutate(across(all_of(vars_nfire), factor))
})

fire_box <- function(xvar, df, method) {
  ggplot(df, aes_string(x= xvar, y = 'value')) +
    geom_boxplot() +
    facet_wrap(~name, scales = 'free_y') +
    labs(subtitle = paste0(xvar, " (", method, " method)"))
}

map2(dfs_biome_long, names(dfs_biome_long), function(df, method) {
  map(vars_nfire, fire_box, df = df, method = method)
})

  
```

## Dotplots: deciles vs fire probability

First calculating mean of response variables (proportion of years with fires) 
for each decile of each predictor
variable. 

Next plotting deciles (originally was deciles, 10 bins, now
actually 0.5 percentiles, i.e. 200 bins)
vs fire probability. Dotplots made for each predictor
variable and response variable (as well method by which response variable
was calculated)

```{r observed_deciles}

dfs_deciles1 <- map(dfs_biome_long, longdf2deciles, response_vars = vars_prop)
# df <- dfs_deciles1[[1]]
# yvar = vars_prop[[1]]

map2(dfs_deciles1, names(dfs_deciles1), function(df, method) {
  map(vars_prop, decile_dotplot, df = df, method = method,
      ylab = "fire probability (per year)")
})

```


## Histograms of response vars

Histograms for all fire variables, includes all grid cells from sagebrush biome,
showing fire probability (nfires/nyears)

```{r hists_response, message = FALSE, warning = FALSE}

dfs_prop_long <- map(dfs_biome3, function(df) {
  df %>% 
    select(matches("_prop$")) %>% 
    pivot_longer(cols = everything(),
                 names_to = "variable",
                 values_to = 'value') %>% 
    mutate(probability = value*100)
})

map(method, function(x) {
  ggplot(dfs_prop_long[[x]], aes(probability)) +
  geom_histogram(breaks = seq(-1, 15)) +
  facet_wrap(~variable, scales = "free_y") +
  labs(x = lab_fireProbPerc,
       caption = paste(x, "method used")) +
  coord_cartesian(xlim = c(-0.1, 15))
})

```

## Plot response variables against each other

```{r response_ggpairs, cache=TRUE, dependson='train'}
map(method, function(x) {
  ggpairs(dfs_test[[x]] , title =  paste0('relationships between response variables (',
                           x,  " method used)"),
        columns = vars_nfire,
        lower = list(continuous = GGally::wrap("points", alpha = 0.1, size=0.2)))
})

```

# GLMs 


## fitting models

### fitting various transforms

Creating formulas where each variable on its own is transformed numerous ways
(including formula to fit spline), all other variables are left alone,
that repeated for each variable. So have models with 1 variable transformed,
2 transformed, etc. 

see documentation for glms_iterate_transforms, in the modelling_functions.R
script

```{r iterate_transform, warning = FALSE, dependson='train', cache = TRUE}
set.seed(1234)

# adding an interaction term to help deal with overpredicting fire probability
# at pixils with high afgaGB and high MAP. perentheses around interaction
# term are included so that glms_iterate_transforms doesn't transform
# the interaction term. 
pred_vars_inter <- c(pred_vars, 'afgAGB:MAP' = "(afgAGB:MAP)",
                     "afgAGB:prcpPropSum" = "(afgAGB:prcpPropSum)")

mods_glm_mtbs1 <- glms_iterate_transforms(
  preds = pred_vars_inter, 
  df = dfs_train$paint[ , ],
  response_var = "mtbs_prop", 
  delta_aic = 10)

# not including the last element of the list which is final_formula
mods_glm_mtbs2 <- mods_glm_mtbs1[-length(mods_glm_mtbs1)]

map_dfc(mods_glm_mtbs2, ~ names(.$aic[1:5])) %>% 
  kable(caption = "5 best transformations at each step")
best_aic_by_step <- map_dbl(mods_glm_mtbs2, ~.$aic[1])

best_aic_by_step <- c(step0 = mods_glm_mtbs1$step1$aic[['convert_none']], 
                      best_aic_by_step)
# AIC improvement by step
diff(best_aic_by_step)
plot(y = best_aic_by_step, 
     x = 1:length(best_aic_by_step) - 1,
     ylab = "AIC",
     xlab = "Number of variables transformed")

```

Best transformation each step.

```{r}
var_transformed <- map(mods_glm_mtbs2, function(x) x$var_transformed)
var_transformed
```


### same model for all response variables

Response variables are the proportion of years in which fires occurred.
Using best model formula selected in the previous step

```{r fit_glm1, dependson='iterate_transform'}

best_form <- mods_glm_mtbs1$final_formula %>% 
  str_replace(".+(?=~)", "") # remove everything before ~ (i.e. response var)
forms_glm1 <-  map(paste(vars_prop, best_form), as.formula)

print(best_form)
names(forms_glm1) <- names(vars_occur)
mod_glm1 <- map(method, function(x) {
  # fitting glms for each formula 
  map(forms_glm1, function(form) {
    glm(form, data = dfs_train[[x]], family = 'binomial', weights = mtbs_n)
  })
}) 

# should be the same AIC (i.e. refitting the same model)
AIC(mod_glm1$paint$mtbs)

mod_glm1_flat <- flatten_rename(mod_glm1)
summary(mod_glm1_flat$paint_mtbs)

```

## partial dependence & VIP

```{r pdp_glm, warning = FALSE, fig.width = 8, fig.height=8, cache = TRUE, dependson='fit_glm1'}
# vip::vip(bin_glm1) # variable importance
vip_plots1 <- map2(mod_glm1_flat, names(mod_glm1_flat), function(mod, name) {
   out <- vip::vip(mod) +
    labs(subtitle = name)
})
marrangeGrob(vip_plots1, ncol = 3, nrow = 3)
map(mod_glm1_flat[1], pdp_all_vars, mod_vars = pred_vars, ylab = 'probability',
              train = train_pred)

# pdp for the original data
map(mod_glm1_flat[1], pdp_all_vars, mod_vars = pred_vars, ylab = 'probability',
              train = df_pred)
```


## observed vs. predicted

Predicting on the data
Note now that dfs_train is the whole dataset, i'm predicting on the
'training' dataframe

```{r}

# create prediction for each each model
# (i.e. for each fire proporation variable)
predict_by_response <- function(mods, df) {
  df_out <- df
  for(var in names(vars_prop)) {
    mod <- mods[[var]]
    
    response_name <- paste0(vars_prop[[var]], "_pred")
    df_out[[response_name]] <- predict(mod, df, type = 'response')
  }
  df_out
}

pred_glm1 <- map2(mod_glm1, dfs_train, predict_by_response)

dfs_test_warm <- map(dfs_train, function(df) mutate(df, MAT = MAT + 5))

pred_glm1_warm <- map2(mod_glm1, dfs_test_warm, predict_by_response)
```

```{r message = FALSE}

# long format so can have observed and predicted columns
pred_glm1_long <- map(pred_glm1, function(df) { 
  df %>% 
    select(matches("_prop")) %>% 
    mutate(row = 1:n()) %>% 
    pivot_longer(cols = -row) %>% 
    mutate(data_source = str_extract(name, "^[a-z]+"),
           type = ifelse(str_detect(name, "prop_pred$"), 'predicted', 
                         'observed')) %>% 
    select(-name) %>% 
    pivot_wider(id_cols = c('row', 'data_source'), 
                names_from = 'type',
                values_from = 'value') %>% 
    select(-row)
})

# observed vs predicted ggplots
g <- ggplot(pred_glm1_long[[1]] %>% filter(observed <= 0.1), 
            aes(y = predicted)) +
  facet_wrap(~data_source) +
  labs(subtitle = 'observed and predicted fire probabilities',
       y = "predicted probability")

g +
  geom_boxplot(aes(x = as.factor(as.numeric(as.factor(observed)) - 1)))+
  labs(x = "number of observed fires")

g +
  geom_jitter(aes(x = observed), width = 0.005, alpha = 0.1) +
  geom_smooth(aes(x = observed), method = 'lm') +
  geom_abline(slope = 1) 

```

### Deciles

Binning predictor variables into deciles (actually 0.5 percentiles) and looking at the mean
predicted probability for each decile.

Then predicting on an identical dataset but with warming

```{r}
vars_prop_pred <- paste0(vars_prop, "_pred")
response_vars <- c(vars_prop, vars_prop_pred)

pred_glm1_deciles <- map(pred_glm1, predvars2deciles, 
                         response_vars = response_vars,
                         pred_vars = pred_vars)

pred_glm1_deciles_warm <- map(pred_glm1_warm, predvars2deciles,
                              response_vars = response_vars,
                              pred_vars = pred_vars)

map2(pred_glm1_deciles, names(pred_glm1_deciles), function(df, method) {
  map(vars_prop, decile_dotplot, df = df, method = method, 
      add_predicted = TRUE)
})

# tryng
map2(pred_glm1_deciles_warm, names(pred_glm1_deciles_warm), function(df, method) {
  map(vars_prop, decile_dotplot, df = df, method = method, 
      add_predicted = TRUE, title = "data w/ +5 deg C warming")
})

```

Plot observed vs predicted values for the quantile

```{r}
# long format, so have columns of predicted and observed
# and can facet by response variable
dfs_obs_v_pred <- map(pred_glm1_deciles, function(df) {
  df %>% 
    select(matches("_prop")) %>% 
    mutate(id = 1:nrow(.)) %>% 
    pivot_longer(cols = matches("_prop")) %>% 
    mutate(type = ifelse(str_detect(name, "_pred$"),
                         "predicted", "observed"),
           variable = str_extract(name, "^[a-z]+")) %>% 
    pivot_wider(id_cols = c(id, variable), names_from = 'type')
})     

map2(dfs_obs_v_pred, names(dfs_obs_v_pred), function(df, method) {
  ggplot(df, aes(observed, predicted)) +
    geom_point() +
    facet_wrap(~variable) +
    geom_abline(slope = 1) +
    labs(title = paste(method, "method"),
         subtitle = "fire probability averaged for each decile of each predictor variable",
         caption = "each panel shows observed and predicted values
         of annual fire probability, for a given fire dataset")
})

```

### Deciles Filtered 

Filtered 'Decile' plots of data. These plots show each vegetation variable,
but only based on data that falls into the upper and lower two deciles of
each climate variable. 


```{r glm_deciles_filtered, fig.height = 6}

pred_glm1_deciles_filt <- map(pred_glm1, predvars2deciles, 
                         response_vars = response_vars,
                         pred_vars = pred_vars,
                         filter_var = TRUE)

pred_glm1_deciles_warm <- map(pred_glm1_warm, predvars2deciles,
                              response_vars = response_vars,
                              pred_vars = pred_vars)

map2(pred_glm1_deciles_filt, names(pred_glm1_deciles_filt), function(df, method) {
  map(vars_prop, decile_dotplot_filtered, df = df, method = method)
})

```


# Save output

```{r save_output}
# glm models
mods2save <- mod_glm1_flat
mods2save$formula <- best_form
saveRDS(mods2save, "models/glm_binomial_models_resample_v1.RDS")

```

# session info

Hash of current commit (i.e. to ID the version of the code used)

```{r}
system("git rev-parse HEAD", intern=TRUE)
```

Packages etc.

```{r}
sessionInfo()

```




