---
title: "Models of fire probability for the sagebrush biome"
subtitle: "Annual (3 year running average) predictor data"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

This dataset consists of annual fire occurrence per cell (i.e. 1 or 0)
and the 3 year average (that year and lag1 and lag2) climate and biomass
of annuals. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, cache.lazy = FALSE,
                      message = FALSE)
```

# Dependencies 

User defined parameters

```{r}
# set to true if want to run for a limited number of rows (i.e. for code testing)
test_run <- FALSE
# _ann defines this model as being built on annual data
s <- "_ann_A-P_A2-T2_A-Pr" # string to paste to file names e.g., that defines the interactions in the model
# such as (summer ppt * MAT) and annuals*temperature interactions
fit_sample <- TRUE # fit model to a sample of the data
n_train <- 5e6

```


```{r warning=FALSE, message=FALSE}
# set option so resampled dataset created here reproduces earlier runs of this code with dplyr 1.0.10
source("src/general_functions.R")
source("src/fig_params.R")
source("src/modeling_functions.R")
source("src/resample_functions.R")
library(tidyverse)
library(GGally) # for ggpairs()
library(pdp) # for partial dependence plots
library(gridExtra)
library(knitr)
library(patchwork) # for figure insets etc. 
theme_set(theme_classic())
```

# read in data

Data compiled in the `04_create_biom-mask_dataframe_ann.R` script

```{r}
df_ann1 <- read_csv('data_processed/fire-clim-veg_3yrAvg_v1.csv',
                    progress = FALSE, show_col_types = FALSE)
```


# Prep data

```{r}
set.seed(1)
df_biome3 <- df_ann1

# small dataset for if testing the data
if(test_run) {
  df_biome3 <- slice_sample(df_biome3, n = 1e5)
}
```


For now, not doing any resampling

```{r}
set.seed(1234)
pred_vars <- c("afgAGB", "pfgAGB", "MAT", "MAP", "prcpPropSum")
names(pred_vars) <- pred_vars
# so that herbAGB is in the final dataframes
pred_vars2 <- pred_vars

# predictor vars are the same in both dfs
df_pred <- df_biome3[, pred_vars2]
```

Training data

```{r}
df_sample <- if(fit_sample) {
  slice_sample(df_biome3, n = n_train)
} else {
    df_biome3
}

# small sample for certain plots
df_small <- slice_sample(df_biome3, n = 1e5)
```


# Exploratory figs & summary values

```{r summary_table}
create_summary <- function(df) {
  df %>% 
    pivot_longer(cols = everything(),
                 names_to = 'variable') %>% 
    group_by(variable) %>% 
    summarise(across(value, .fns = list(mean = mean, min = min, 
                                        median = median, max = max))) %>% 
    mutate(across(where(is.numeric), round, 4))
}

df_biome3[pred_vars2] %>% 
  create_summary() %>% 
  knitr::kable(caption = 'summaries of predictor variables')


response_summary <- df_biome3 %>% 
    dplyr::select(where(is.numeric), -all_of(pred_vars2), -matches("^occur_")) %>% 
    create_summary()


kable(response_summary, 
      caption = 'summaries of response variables, calculated using paint')

```

## Plot predictor vars against each other

here using pred dataframe, because smaller and this code is slow. 

```{r pred_v_pred, cache = TRUE, dependson='train'}
df_pred %>% 
  slice_sample(n = 5e4) %>% 
  select(-matches("_")) %>% 
ggpairs(lower = list(continuous = GGally::wrap("points", alpha = 0.1,    size=0.2)))

```

## boxplots-- # of fires vs predictor variables

```{r n_fires_boxplots, fig.height=9, fig.width=8, cache = FALSE, dependson='prep_data'}

# vectors of names of response variables
vars_nfire <- names(df_biome3) %>% 
  str_subset("^nfire_")

var_prop <- 'cwf_prop'

names(var_prop) <- 'cwf'

# longformat dataframes for making boxplots
df_biome_long <- 
  # for some reason select() was giving me problems
  # adding numYrs here so can take weighted average
  predvars2long(df_biome3, response_vars = c(vars_nfire, var_prop), 
                pred_vars = pred_vars) %>% 
    mutate(across(all_of(vars_nfire), factor))


ggplot(df_biome_long, aes_string(x= "nfire_cwf", y = 'value')) +
  geom_boxplot() +
  facet_wrap(~name, scales = 'free_y')

```


# GLMs 

## fitting models

### fitting various transforms

Creating formulas where each variable on its own is transformed numerous ways
(including formula to fit spline), all other variables are left alone,
that repeated for each variable. So have models with 1 variable transformed,
2 transformed, etc. 

see documentation for glms_iterate_transforms, in the modelling_functions.R
script

```{r iterate_transform, warning = FALSE, dependson='train', cache = FALSE}
set.seed(1234)

# adding an interaction term to help deal with over-predicting fire probability
# at pixels with high afgAGB and high MAP. parentheses around interaction
# term are included so that glms_iterate_transforms doesn't transform
# the interaction term. 
pred_vars_inter <- c(pred_vars, 'afgAGB:MAP' = "(afgAGB:MAP)",
                     #"afgAGB:prcpPropSum" = "(afgAGB:prcpPropSum)",
                     # based on examination of diagnostics plots with 
                     # Daniel, decided to try fitting this interaction
                     # (annuals and temp) as a parabola to help avoid
                     # over-predictions in warm high annual places. 
                     "afgAGB:MAT" = "(afgAGB:MAT)",
                     "afgAGB2:MAT2" = "I(afgAGB^2):I(MAT^2)",
                     #"afgAGB:prcpPropSum" = "(afgAGB:prcpPropSum)",
                     #"afgAGB2:prcpPropSum2" = "I(afgAGB^2):I(prcpPropSum^2)",
                     "afgAGB:pfgAGB" = "(afgAGB:pfgAGB)"#,
                     #"afgAGB2:pfgAGB2" = "I(afgAGB^2):I(pfgAGB^2)"
                     )

# pred_vars_inter <- pred_vars

mods_glm_cwf1 <- glms_iterate_transforms(
  preds = pred_vars_inter, 
  df = df_sample,
  response_var = "cwf_prop", 
  delta_aic = 10,
  weights_col = "numYrs")

# not including the last element of the list which is final_formula
mods_glm_cwf2 <- mods_glm_cwf1[-length(mods_glm_cwf1)]

map_dfc(mods_glm_cwf2, ~ names(.$aic[1:5])) %>% 
  kable(caption = "5 best transformations at each step")
best_aic_by_step <- map_dbl(mods_glm_cwf2, ~.$aic[1])

best_aic_by_step <- c(step0 = mods_glm_cwf1$step1$aic[['convert_none']], 
                      best_aic_by_step)
# AIC improvement by step
diff(best_aic_by_step)
plot(y = best_aic_by_step, 
     x = 1:length(best_aic_by_step) - 1,
     ylab = "AIC",
     xlab = "Number of variables transformed")

```

Best transformation each step.

```{r}
var_transformed <- map(mods_glm_cwf2, function(x) x$var_transformed)
var_transformed
```


### same model for all response variables

Response variables are the proportion of years in which fires occurred.
Using best model formula selected in the previous step

```{r fit_glm1, dependson='iterate_transform'}

best_form <- as.formula(mods_glm_cwf1$final_formula)
print(best_form)

# refitting the glm with the best formula
mod_glm1 <- glm(best_form, data = df_sample, family = 'binomial')

# should be the same AIC (i.e. refitting the same model)
AIC(mod_glm1)

summary(mod_glm1)

```

## partial dependence & VIP

PDP plot trend made using a small sample of the data

```{r pdp_glm, warning = FALSE, fig.width = 8, fig.height=8, cache = FALSE, dependson='fit_glm1'}
# vip::vip(bin_glm1) # variable importance
vip::vip(mod_glm1, num_features = 15)

pdp_all_vars(mod_glm1, mod_vars = pred_vars, ylab = 'probability',train = df_small)

```


## observed vs. predicted

Predicting on the data

```{r}

# create prediction for each each model
# (i.e. for each fire proporation variable)
predict_by_response <- function(mod, df) {
  df_out <- df

  response_name <- paste0(var_prop, "_pred")
  df_out[[response_name]] <- predict(mod, df, type = 'response')
  df_out
}

pred_glm1 <- predict_by_response(mod_glm1, df_biome3)
```

```{r message = FALSE}

# long format so can have observed and predicted columns
pred_glm1_long <- pred_glm1 %>% 
    select(matches("_prop")) %>% 
    mutate(row = 1:n()) %>% 
    pivot_longer(cols = -row) %>% 
    mutate(data_source = str_extract(name, "^[a-z]+"),
           type = ifelse(str_detect(name, "prop_pred$"), 'predicted', 
                         'observed')) %>% 
    select(-name) %>% 
    pivot_wider(id_cols = c('row', 'data_source'), 
                names_from = 'type',
                values_from = 'value') %>% 
    select(-row)

# observed vs predicted ggplots
g <- ggplot(pred_glm1_long %>% filter(observed <= 0.1), 
            aes(y = predicted)) +
  facet_wrap(~data_source) +
  labs(subtitle = 'observed and predicted fire probabilities',
       y = "predicted probability")

g +
  geom_boxplot(aes(x = as.factor(as.numeric(as.factor(observed)) - 1)))+
  labs(x = "number of observed fires")


```

### Deciles

Binning predictor variables into deciles (actually percentiles) and looking at the mean
predicted probability for each percentile. The use of the word deciles
is just a legacy thing (they started out being actual deciles)

Then predicting on an identical dataset but with warming

```{r}
var_prop_pred <- paste0(var_prop, "_pred")
response_vars <- c(var_prop, var_prop_pred)

pred_glm1_deciles <- predvars2deciles(pred_glm1,
                                      response_vars = response_vars,
                                      pred_vars = pred_vars)

```



Publication quality quantile plot

```{r}

# publication quality version
g <- decile_dotplot_pq(pred_glm1_deciles)

# obs/pred inset
g2 <- add_dotplot_inset(g, pred_glm1_deciles)

g2
png(paste0("figures/quantile_plots/quantile_plot_v5", s,  ".png"), 
     units = "in", res = 600, width = 5.5, height = 3.5 )
g2
dev.off()
```

### Deciles Filtered 

20th and 80th percentiles for each climat variable

```{r}
df <- pred_glm1[, c("MAT", "MAP", "prcpPropSum")] %>% 
  mutate(MAT = MAT - 273.15) # k to c
map(df, quantile, probs = c(0.2, 0.8), na.rm = TRUE)
```


Filtered 'Decile' plots of data. These plots show each vegetation variable,
but only based on data that falls into the upper and lower two deciles of
each climate variable. 


```{r glm_deciles_filtered, fig.height = 10, fig.width = 5, message = FALSE}
clim_vars <- c("MAT", "MAP", "prcpPropSum")
pred_glm1_deciles_filt <- predvars2deciles( pred_glm1, 
                         response_vars = response_vars,
                         pred_vars = pred_vars,
                         filter_var = TRUE,
                         filter_vars = pred_vars) 

decile_dotplot_filtered_pq(pred_glm1_deciles_filt, xvars = clim_vars)
decile_dotplot_filtered_pq(pred_glm1_deciles_filt)

```



Filtered quantile figure with middle 2 deciles also shown

```{r fig.height = 8, fig.width = 5}
pred_glm1_deciles_filt_mid <- predvars2deciles(pred_glm1, 
                         response_vars = response_vars,
                         pred_vars = pred_vars,
                         filter_var = TRUE,
                         add_mid = TRUE)

g <- decile_dotplot_filtered_pq(df = pred_glm1_deciles_filt_mid)
g

jpeg(paste0("figures/quantile_plots/quantile_plot_filtered_mid_v1", s, ".jpeg"),
     units = "in", res = 600, width = 5.5, height = 6 )
g 
dev.off()
```



# Save output

```{r save_output}
# glm models
mods2save <- butcher::butcher(mod_glm1) # removes some model components so the saved object isn't huge
mods2save$summary <- summary(mod_glm1) # summary doesn't work on the butchered object
mods2save$formula <- best_form
mods2save$pred_vars_inter <- pred_vars_inter # so have interactions
n <- nrow(df_sample)
mods2save$data_rows <- n

if(!test_run) {
  saveRDS(mods2save, 
        paste0("models/glm_binomial_models_v1", s, "_", n, "n.RDS"))
}


```

# session info

Hash of current commit (i.e. to ID the version of the code used)

```{r}
system("git rev-parse HEAD", intern=TRUE)
```

Packages etc.

```{r}
sessionInfo()

```




