---
title: "GAM of fire probability for the sagebrush biome"
subtitle: "Annual (3 year running average) predictor data"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

This dataset consists of annual fire occurrence per cell (i.e. 1 or 0)
and the 3 year average (that year and lag1 and lag2) climate and biomass
of annuals. 

Fitting GAM with no interactions, to better visualize what the non linear
relationship would look like if a more flexible model were used. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, cache.lazy = FALSE,
                      message = FALSE)
```

# Dependencies 

User defined parameters

```{r}
# set to true if want to run for a limited number of rows (i.e. for code testing)
test_run <- FALSE
# _ann defines this model as being built on annual data
s <- "_ann_gam" # string to paste to file names e.g., that defines the interactions in the model
# such as (summer ppt * MAT) and annuals*temperature interactions
fit_sample <- TRUE # fit model to a sample of the data
n_train <- 2e6
n_test <- 1e6 # the q

```


```{r warning=FALSE, message=FALSE}
# set option so resampled dataset created here reproduces earlier runs of this code with dplyr 1.0.10
source("src/general_functions.R")
source("src/fig_params.R")
source("src/modeling_functions.R")
source("src/resample_functions.R")
library(tidyverse)
library(GGally) # for ggpairs()
library(pdp) # for partial dependence plots
library(gridExtra)
library(knitr)
library(patchwork) # for figure insets etc. 
theme_set(theme_classic())
```

# read in data

Data compiled in the `04_create_biom-mask_dataframe_ann.R` script

```{r}
df_ann1 <- read_csv('data_processed/fire-clim-veg_3yrAvg_v1.csv',
                    progress = FALSE, show_col_types = FALSE)
```


# Prep data

```{r}
set.seed(1)
df_biome3 <- df_ann1

# small dataset for if testing the data
if(test_run) {
  df_biome3 <- slice_sample(df_biome3, n = 1e5)
}
```


For now, not doing any resampling

```{r}
set.seed(1234)
pred_vars <- c("afgAGB", "pfgAGB", "MAT", "MAP", "prcpPropSum")
names(pred_vars) <- pred_vars
# so that herbAGB is in the final dataframes
pred_vars2 <- pred_vars

# predictor vars are the same in both dfs
df_pred <- df_biome3[, pred_vars2]
```

Training data

```{r}
df_sample <- if(fit_sample) {
  slice_sample(df_biome3, n = n_train)
} else {
    df_biome3
}

df_test <- if(fit_sample & !test_run & 
              # antijoin only works if there are enough rows that meet 
              # that criterion, i.e. if df_sample contains most of the data i
              # doesnt' work
              (nrow(df_biome3) - nrow(df_sample) > n_test)) {
  df_biome3 %>% 
    anti_join(df_sample, by = c("cell_num", "year")) %>% 
    slice_sample(n = n_test)
} else {
  df_biome3 %>% 
    slice_sample(n = n_test)
}

# small sample for certain plots
df_small <- slice_sample(df_biome3, n = 1e5)
```

# describe data

Number of fires in the data used to fit the model

```{r}
sum(df_sample$nfire_cwf)
```

# GAMS 

## fitting models


```{r fit_gam1}

# refitting the gam with the best formula
best_form <- formula(cwf_prop ~ s(afgAGB, bs = 'bs') + s(pfgAGB, bs = 'bs') + 
                  s(MAT, bs = 'bs') + s(MAP, bs = 'bs') + s(prcpPropSum, bs = 'bs'))
mod_gam1 <- mgcv::bam(best_form,
                data = df_sample, family = 'binomial')

# should be the same AIC (i.e. refitting the same model)

summary(mod_gam1)

```

## partial dependence & VIP

PDP plot trend made using a small sample of the data

```{r pdp_gam, warning = FALSE, fig.width = 8, fig.height=8}
# vip::vip(bin_gam1) # variable importance

pdp_all_vars(mod_gam1, mod_vars = pred_vars, ylab = 'probability',train = df_small)

```


## observed vs. predicted

Predicting on the data

```{r}
var_prop <- 'cwf_prop'
# create prediction for each each model
# (i.e. for each fire proporation variable)
predict_by_response <- function(mod, df) {
  df_out <- df

  response_name <- paste0(var_prop, "_pred")
  df_out[[response_name]] <- predict(mod, df, type = 'response')
  df_out
}

pred_gam1 <- predict_by_response(mod_gam1, df_biome3)
```

```{r message = FALSE}

# long format so can have observed and predicted columns
pred_gam1_long <- pred_gam1 %>% 
    select(matches("_prop")) %>% 
    mutate(row = 1:n()) %>% 
    pivot_longer(cols = -row) %>% 
    mutate(data_source = str_extract(name, "^[a-z]+"),
           type = ifelse(str_detect(name, "prop_pred$"), 'predicted', 
                         'observed')) %>% 
    select(-name) %>% 
    pivot_wider(id_cols = c('row', 'data_source'), 
                names_from = 'type',
                values_from = 'value') %>% 
    select(-row)

# observed vs predicted ggplots
g <- ggplot(pred_gam1_long %>% filter(observed <= 0.1), 
            aes(y = predicted)) +
  facet_wrap(~data_source) +
  labs(subtitle = 'observed and predicted fire probabilities',
       y = "predicted probability")

g +
  geom_boxplot(aes(x = as.factor(as.numeric(as.factor(observed)) - 1)))+
  labs(x = "number of observed fires")


```

### Deciles

Binning predictor variables into deciles (actually percentiles) and looking at the mean
predicted probability for each percentile. The use of the word deciles
is just a legacy thing (they started out being actual deciles)

Then predicting on an identical dataset but with warming

```{r}
var_prop_pred <- paste0(var_prop, "_pred")
response_vars <- c(var_prop, var_prop_pred)

pred_gam1_deciles <- predvars2deciles(pred_gam1,
                                      response_vars = response_vars,
                                      pred_vars = pred_vars)

```



Publication quality quantile plot

```{r}

# publication quality version
g <- decile_dotplot_pq(pred_gam1_deciles)

# obs/pred inset
g2 <- add_dotplot_inset(g, pred_gam1_deciles)

g2
if(!test_run){
png(paste0("figures/quantile_plots/quantile_plot_v5", s,  ".png"), 
     units = "in", res = 600, width = 5.5, height = 3.5 )
g2
dev.off()
}
```

### Deciles Filtered 

20th and 80th percentiles for each climat variable

```{r}
df <- pred_gam1[, c("MAT", "MAP", "prcpPropSum")] %>% 
  mutate(MAT = MAT - 273.15) # k to c
map(df, quantile, probs = c(0.2, 0.8), na.rm = TRUE)
```


Filtered 'Decile' plots of data. These plots show each vegetation variable,
but only based on data that falls into the upper and lower two deciles of
each climate variable. 


```{r gam_deciles_filtered, fig.height = 10, fig.width = 5, message = FALSE}
clim_vars <- c("MAT", "MAP", "prcpPropSum")
pred_gam1_deciles_filt <- predvars2deciles( pred_gam1, 
                         response_vars = response_vars,
                         pred_vars = pred_vars,
                         filter_var = TRUE,
                         filter_vars = pred_vars) 

decile_dotplot_filtered_pq(pred_gam1_deciles_filt, xvars = clim_vars)
decile_dotplot_filtered_pq(pred_gam1_deciles_filt)

```


# Save output

```{r save_output}
# gam models
mods2save <- butcher::butcher(mod_gam1) # removes some model components so the saved object isn't huge
mods2save$summary <- summary(mod_gam1) # summary doesn't work on the butchered object
mods2save$formula <- best_form
n <- nrow(df_sample)
mods2save$data_rows <- n

if(!test_run) {
  saveRDS(mods2save, 
        paste0("models/gam_binomial_models_v1", s, "_", n, "n.RDS"))
}


```

# session info

Hash of current commit (i.e. to ID the version of the code used)

```{r}
system("git rev-parse HEAD", intern=TRUE)
```

Packages etc.

```{r}
sessionInfo()

```




