---
title: "Models of fire probability for the sagebrush biome"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

# Dependencies 

Sourcing the script that creates a dataframe of historical
fore occurrence, as well as biotic and climate
predictors for those same grid-cells. 

```{r warning=FALSE, message=FALSE, cache = TRUE}
source("scripts/04_create_biome-mask_dataframe.R")
```

```{r warning=FALSE, message=FALSE}
source("src/general_functions.R")
source("src/fig_params.R")
source("src/modeling_functions.R")
library(randomForest)
library(tidyverse)
library(margins) # for cplot()
library(GGally) # for ggpairs()
library(gnm) # generalized non-linear models
library(pdp) # for partial dependence plots
library(gridExtra)
library(knitr)
theme_set(theme_classic())
```


# Prep data

```{r prep_data}

# converting a fire count into a factor (true if at least 1 fire occured)
occur_factor <- function(nfire){
  factor(nfire > 0, levels = c("TRUE", "FALSE"))
}
dfs_biome3 <- map(dfs_biome2, function(df) {
  out <- df %>% 
  # probability of fire in a given year
    mutate(# number of years the mtbs count data corresponds to (for binomial glm)
           # all count datasets (mtbs and ifph) are 35 years (change if updated)
           mtbs_n = 36, 
           # proportion of years with fires
           mtbs_prop = nfire_mtbs/mtbs_n,
           ifph_prop = nfire_ifph/mtbs_n,
           comb_prop = nfire_comb/mtbs_n,
           lba_prop = nfire_lba/mtbs_n) %>% 
    # creating true/false occurrence cols for each datasets
    mutate(across(matches("^nfire_"), .fns = list(occur = occur_factor),
                  .names = "{.fn}_{.col}")) %>% 
    rename_with(.fn = str_replace, .cols = matches("^occur_"),
                pattern = "_nfire", replacement = "")
})

method <- names(dfs_biome3)
names(method) <- method
```

Create transformed variables

```{r}
dfs_biome3 <- map(dfs_biome3, function(df) {
  df %>% 
    mutate(afgAGB_sqrt = sqrt(afgAGB),
           afgAGB_ln = log(afgAGB),
           prcpPropSum_exp = exp(prcpPropSum),
           prcpPropSum_sqrt = sqrt(prcpPropSum),
           prcpPropSum_sq = prcpPropSum^2,
           MAP_sqrt = sqrt(MAP))
})
```



## training data

Using a small sample for now to make model fitting quicker

```{r train}
set.seed(1234)
pred_vars <- c("afgAGB", "pfgAGB", "shrCover", "MAT", "MAP", "prcpPropSum")

# predictor vars are the same in both dfs
df_pred <- dfs_biome3[[1]][, pred_vars]
n = 10000
row_nums <- 1:nrow(dfs_biome3[[1]])
rows_train <- sample(row_nums, size = n, replace = FALSE)

# rows for test dataset
rows_test <- sample(row_nums[!row_nums %in% rows_train], size = 40000, replace = FALSE)

# selecting rows this was so both dataframes in list have same rows
# and differences in models will only be b/ of differences in response
dfs_train <- map(dfs_biome3, function(df) as.data.frame(df[rows_train, ]))
dfs_test <- map(dfs_biome3, function(df) as.data.frame(df[rows_test, ]))

# all variables (transformed or not, that could be predictor vars)
all_pred_vars <- names(dfs_biome3[[1]]) %>% 
  str_subset(pattern = paste0("(", 
                              paste0(pred_vars, collapse = ")|("), 
                              ")")
             )
train_pred <- dfs_train[[1]][, all_pred_vars]
test_pred <- dfs_test[[1]][, all_pred_vars]
```

# Exploratory figs & summary values


```{r}

create_summary <- function(df) {
  df %>% 
    pivot_longer(cols = everything(),
                 names_to = 'variable') %>% 
    group_by(variable) %>% 
    summarise(across(value, .fns = list(mean = mean, min = min, 
                                        median = median, max = max))) %>% 
    mutate(across(where(is.numeric), round, 4))
}

df_pred %>% 
  create_summary() %>% 
  knitr::kable(caption = 'summaries of predictor variables')


response_summary <- map(dfs_biome3, function(df) {
  df %>% 
    dplyr::select(-all_of(pred_vars), -matches("^occur_")) %>% 
    create_summary()
})

kable(response_summary$paint, 
      caption = 'summaries of response variables, calculated using paint  <- ')

kable(response_summary$reduceToImage, 
      caption = 'summaries of response variables, calculated using reduceToImage method')

```

## Plot predictor vars against each other

```{r}
ggpairs(train_pred,
        lower = list(continuous = GGally::wrap("points", alpha = 0.1,    size=0.2)))
```

## boxplots--predictor variables vs. # of fires

```{r fig.height=9, fig.width=8, cache = TRUE, dependson='prep_data'}

vars_nfire <- names(dfs_biome3[[1]]) %>% 
  str_subset("^nfire_")
names(vars_nfire) <- vars_nfire
df <- dfs_biome3[[1]] %>% 
  pivot_longer(cols = all_of(pred_vars)) %>% 
  mutate(across(all_of(vars_nfire), factor))

# dataframes for making boxplots
dfs_biome_long <- map(dfs_biome3, function(df) {
  df %>% 
    pivot_longer(cols = all_of(pred_vars)) %>% 
    mutate(across(all_of(vars_nfire), factor))
})

fire_box <- function(xvar, df, method) {
  ggplot(df, aes_string(x= xvar, y = 'value')) +
    geom_boxplot() +
    facet_wrap(~name, scales = 'free_y') +
    labs(subtitle = paste0(xvar, " (", method, " method)"))
}

map2(dfs_biome_long, names(dfs_biome_long), function(df, method) {
  map(vars_nfire, fire_box, df = df, method = method)
})

  

```

## Histograms of response vars

Histograms for all fire variables, includes all grid cells from sagebrush biome,
showing fire probability (nfires/nyears)

```{r message = FALSE, warning = FALSE}

dfs_prop_long <- map(dfs_biome3, function(df) {
  df %>% 
    select(matches("_prop$")) %>% 
    pivot_longer(cols = everything(),
                 names_to = "variable",
                 values_to = 'value') %>% 
    mutate(probability = value*100)
})

map(method, function(x) {
  ggplot(dfs_prop_long[[x]], aes(probability)) +
  geom_histogram(breaks = seq(-1, 15)) +
  facet_wrap(~variable, scales = "free_y") +
  labs(x = lab_fireProbPerc,
       caption = paste(x, "method used")) +
  coord_cartesian(xlim = c(-0.1, 15))
})

```

## Plot response variables against each other

```{r}
map(method, function(x) {
  ggpairs(dfs_train[[x]] , title =  paste0('relationships between response variables (',
                           x,  " method used)"),
        columns = vars_nfire,
        lower = list(continuous = GGally::wrap("points", alpha = 0.1, size=0.2)))
})

```

# Random forest


## Fit Model

```{r}
nodesize <-  50 # minimum size of terminal nodes in trees
vars_occur <- names(dfs_biome3[[1]]) %>% 
  str_subset("^occur_")

names(vars_occur) <- str_replace(vars_occur, "occur_", "")
# fire occurence (0/1) in MTBS data, over 35 years
forms_occur <-  map(paste(vars_occur, ' ~', paste(pred_vars, collapse = "+")),
                    as.formula)

names(forms_occur) <- names(vars_occur)

# fitting random forests to dfs of response variables created with both methods
rf_mods1 <- map(method, function(x) {
  # fitting rfs for each formula 
  map(forms_occur, randomForest, ntree = 100, data = dfs_train[[x]],
               nodesize = nodesize)
})
rf_mods2 <- flatten_rename(rf_mods1)

```

## Variable importance

```{r results='hide', fig.keep='all'}

par(mfrow = c(2, 2))
map2(rf_mods2, names(rf_mods2), function(x, y) {
  print(varImpPlot(x = x, main = y))
})

```

## Partial dependence plots

The output is transformed to yearly fire probability. Rugs on plots show
the percentiles of the predictor variable

```{r fig.height=5, fig.width = 8, cache = TRUE}

pdp_all_rf_mods(mod_list = rf_mods2, df_train = train_pred)

```


## Examine Model fit

The strongly bi-modal histogram suggests this model is overfitting (i.e.
 probabilities close to 1 and 0)

```{r}
mod <- rf_mods2$paint_mtbs
y <- predict(mod, test_pred, type = "prob")
hist(calc_yearly_prob(y), main = "Predicted probabilities (per year)")
```

## transforming predictor variables

### Fit Model

```{r}
pred_vars2 <- c("afgAGB_sqrt", "pfgAGB", "shrCover", "MAP", "MAT", 
                "prcpPropSum_sqrt")

# fire occurence (0/1) in MTBS data, over 35 years
forms_occur2 <-  map(paste(vars_occur, ' ~', paste(pred_vars2, collapse = "+")),
                    as.formula)

names(forms_occur2) <- str_replace(vars_occur, "occur_", "")
# fitting random forests to dfs of response variables created with both methods
rf_mods3 <- map(method, function(x) {
  # fitting rfs for each formula 
  map(forms_occur2, randomForest, ntree = 100, data = dfs_train[[x]],
               nodesize = nodesize)
}) %>% 
  flatten_rename()

```


### Partial dependence plots

The output is transformed to yearly fire probability. Rugs on plots show
the percentiles of the predictor variable

```{r fig.height=5, fig.width = 8, cache = TRUE}

pdp_all_rf_mods(mod_list = rf_mods3[1], df_train = train_pred)

```

# GLMs

## fitting models

Response variables are the proportion of years in which fires occurred

```{r}

vars_prop <- str_replace(vars_occur, "occur_", "") %>% 
  paste0(., "_prop")
names(vars_prop) <- names(vars_occur)

forms_glm1 <-  map(paste(vars_prop, "~ sqrt(afgAGB) + pfgAGB + shrCover", 
                         "+ MAP + MAT + sqrt(prcpPropSum)"), as.formula)

names(forms_glm1) <- names(vars_occur)
mod_glm1 <- map(method, function(x) {
  # fitting glms for each formula 
  map(forms_glm1, function(form) {
    glm(form, data = dfs_train[[x]], family = 'binomial', weights = mtbs_n)
  })
}) 

mod_glm1_flat <- flatten_rename(mod_glm1)
map(mod_glm1_flat, summary)

```

## partial dependence & VIP


```{r, warning = FALSE, fig.width = 8, fig.height=8, cache = TRUE}
# vip::vip(bin_glm1) # variable importance
vip_plots1 <- map2(mod_glm1_flat, names(mod_glm1_flat), function(mod, name) {
   out <- vip::vip(mod) +
    labs(subtitle = name)
})
marrangeGrob(vip_plots1, ncol = 3, nrow = 3)
map(mod_glm1_flat, pdp_all_vars, mod_vars = pred_vars, ylab = 'probability',
              train = df_pred)
```


## observed vs. predicted

Predicting on different rows than models fit using

```{r}
pred_glm1 <- map2(mod_glm1, dfs_train, function(mods, df) {
  df_out <- df
  for(var in names(vars_prop)) {
    mod <- mods[[var]]
    
    response_name <- paste0(vars_prop[[var]], "_pred")
    df_out[[response_name]] <- predict(mod, df, type = 'response')
  }
  df_out
})
```

```{r message = FALSE}

# long format so can have observed and predicted columns
pred_glm1_long <- map(pred_glm1, function(df) { 
  df %>% 
    select(matches("_prop")) %>% 
    mutate(row = 1:n()) %>% 
    pivot_longer(cols = -row) %>% 
    mutate(data_source = str_extract(name, "^[a-z]+"),
           type = ifelse(str_detect(name, "prop_pred$"), 'predicted', 
                         'observed')) %>% 
    select(-name) %>% 
    pivot_wider(id_cols = c('row', 'data_source'), 
                names_from = 'type',
                values_from = 'value') %>% 
    select(-row)
})

# observed vs predicted ggplots
g <- ggplot(pred_glm1_long[[1]] %>% filter(observed <= 0.1), 
            aes(y = predicted)) +
  facet_wrap(~data_source) +
  labs(subtitle = 'observed and predicted fire probabilities',
       y = "predicted probability")

g +
  geom_boxplot(aes(x = as.factor(as.numeric(as.factor(observed)) - 1)))+
  labs(x = "number of observed fires")

g +
  geom_jitter(aes(x = observed), width = 0.005, alpha = 0.1) +
  geom_smooth(aes(x = observed), method = 'lm') +
  geom_abline(slope = 1) 

```

