---
title: "Resampling observed fire probability data from the sagebrush biome"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

Here the training data is resampled to be more balanced (i.e., upsampling
bins of rare data, downsampling bins of common data). 

Note a concern with the up-sampling is that many bins start with only one
datapoint in them, which is then repeated many times in the resampled datset.
Apparently, having duplicated points like this can increase the risk of 
over-fitting. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache.lazy = FALSE,
                      message = FALSE)
```

# Dependencies 

Sourcing the script that creates a dataframe of historical
fore occurrence, as well as biotic and climate
predictors for those same grid-cells. 

```{r source_data, warning=FALSE, message=FALSE}
source("scripts/04_create_biome-mask_dataframe.R")
```

```{r warning=FALSE, message=FALSE}
source("src/general_functions.R")
source("src/modeling_functions.R")
source("src/resample_functions.R")
theme_set(theme_classic())
```


# Prep data

```{r prep_data}

method <- names(dfs_biome3)
names(method) <- method

# dfs_biome3 <- map(dfs_biome3, slice_sample, n = 1000) # for testing
```

```{r}
set.seed(1234)
pred_vars <- c("afgAGB", "pfgAGB", "MAT", "MAP", "prcpPropSum")
names(pred_vars) <- pred_vars
# so that herbAGB is in the final dataframes
pred_vars2 <- c("afgAGB", "pfgAGB", 'herbAGB', "MAT", "MAP", "prcpPropSum")
names(pred_vars2) <- pred_vars2
# predictor vars are the same in both dfs
df_pred <- dfs_biome3[[1]][, pred_vars2]
```


## Bin data

```{r}
dfs_bin <- list()
bin_values <- c(5, 10, 20, 29)
bin_names <- paste0("bin", bin_values)

names(bin_values) <- bin_names
dfs_bin <- map(bin_values, function(x) {
  bin_df(dfs_biome3$paint, cols = pred_vars,
                        n_categories = x)
})

```

## Explore bins

Number of unique bins

```{r}
n_bins <- map_dbl(dfs_bin, function(df){ 
  df$bin_all %>% 
    unique %>% 
    length()
  })
n_bins

```

Percent of total possible number of bins filled

```{r}
(n_bins/(bin_values^5)*100) %>% 
  round(1)
```


Number of observations per bin

```{r}
n_per_bin <- map(dfs_bin, function(df) {
  as.vector(table(df$bin_all))
})



map(n_per_bin, summary)
```

Percent of bins with 1 observation

```{r}
map_dbl(n_per_bin, function(x) mean(x ==1)*100)

```

## Resample data

Re-sampling multiple times, to get a sense for the variability
between random draws

```{r resample, message = FALSE}

samples <- letters[1:3]

dfs_resampled <- list()

for(x in samples) {
  dfs_resampled[[x]] <- map(dfs_bin, resample_bins)
}

dfs_resampled <- flatten_rename(dfs_resampled)

```


# Quantile plots

Calculating mean values per percentile of each predictor variable, for both
the original data and the resampled data

```{r}

response_vars <- "mtbs_prop"
deciles_obs <- predvars2deciles(dfs_biome3$paint, 
                               response_vars = response_vars,
                               pred_vars = pred_vars) %>% 
  mutate(resample = "Original data")

deciles_resampled1 <- map_dfr(dfs_resampled, predvars2deciles,
                         response_vars = response_vars,
                         pred_vars = pred_vars,
                         .id = 'resample')

deciles_comb <- bind_rows(deciles_obs,
                          deciles_resampled1) %>% 
  mutate(replicate = str_extract(resample, "^[a-z](?=_)"),
         type = str_replace(resample, "^[a-z]_", ""))
deciles_comb %>% 
  arrange(mean_value)

```

creating the plots.

Note multiple lines per type, shows the multiple random samplings,
which are all almost identical. 

```{r, fig.height=8, fig.width = 9}

# colors--I don't like the lightest color, so not using it
cols <- RColorBrewer::brewer.pal(length(dfs_bin) + 1, "YlOrBr")[-1]
names(cols) <- names(dfs_bin)
cols <- c("Original data" = "black",
          cols)

decile_base <- function(x) {
  list(    
    facet_wrap(~name, scales = 'free_x'),
  labs(x = "mean of quantile of predictor variable",
       y = "Observed fire probability (mtbs)",
       subtitle = "Original and resampled data",
       caption = "bin(n) refers to the number of bins each
       predictor variable was split into"),
  theme(legend.position = 'top',
        legend.title = element_blank()),
  scale_color_manual(
    values = cols))

}

g <- ggplot(deciles_comb, aes(x = mean_value, y = mtbs_prop,
            color = type, group = resample)) +
   geom_smooth(se = FALSE, size =0.75) +
   decile_base() 
  
g
```

Decile plots showing points for just a single bin size at a time

```{r, fig.height=7, fig.width = 8}

bin_names_a <- paste0("a_", bin_names)
for(x in bin_names_a) {
  print(x)
  df <- deciles_comb %>% 
    filter(resample %in% c("Original data", x))
  
  g <- ggplot(df, aes(x = mean_value, y = mtbs_prop,
            color = type, group = resample)) +
   geom_smooth(se = FALSE, size =0.75) +
    geom_point() +
    decile_base() 
  
  print(g)
  
}

```

# Histograms

Combining into a single dataframe. (only using one random sample of
the bin resampling)

```{r }

# grabbing just the first random sample
df_comb <- dfs_resampled[bin_names_a] %>% 
  # selecting certain rows just resulting dataframe is smaller
  map(function(df) df[pred_vars2]) %>% 
  bind_rows(.id = "resample") %>% 
  mutate(resample = str_replace(resample, "a_", "")) %>% 
  bind_rows(dfs_biome3$paint[pred_vars2] %>% 
              mutate(resample = "Original data")) %>% 
  mutate(resample = factor(resample, levels = names(cols)))

comb_means <- df_comb %>% 
  group_by(resample) %>% 
  summarise(across(all_of(pred_vars2), mean))
  
```

Histograms of each predictor variable, 

```{r histograms, fig.height = 8}
map(pred_vars2, function(x) {
  g <- ggplot(df_comb) +
    geom_histogram(aes_string(x = x), bins = 100) +
    facet_wrap(~resample, ncol = 2, scales = 'free_y') +
    geom_vline(data = comb_means, aes_string(xintercept = x),
               linetype = 2) +
    labs(title = x,
        subtitle = "Histograms of original and resampled data",
         caption = "Vertical line shows the mean value")
  g
})
```


